Long- and short-read metabarcoding technologies reveal similar spatio-temporal structures in fungal communities

### 2020-10-12 writing/transect_paper.Rmd
start addressing mohammad's comments, move some methods to the supplement.

base.dir <- regmatches(getwd(), regexpr(".*oueme-fungi-transect", getwd()))
write_bib(c("base", "phyloseq", "vegan", "metacoder"), file = "R.bib", tweak = TRUE)

@@

Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
Knowledge of turnover scale is import when planning studies of local diversity and its environmental correlates.
It varies between different ecosystems and taxonomic groups.
Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs. distance [@legendre2012].

@@

We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity.
The DADA2 algorithm requires that the seed sequence of each ASV be represented by at least two error-free reads [@callahan2016].
If sequencing errors are uniformly distributed, then the probability that a given read will be error-free is $(1-\epsilon)^L$, where $\epsilon$ is the sequencing error rate and $L$ is the read length in base pairs.
Then the number of reads of a given sequence that would be required to obtain two error-free reads in expectation is $2/(1-\epsilon)^L$.
For the combination of long reads (median $L=`r readd(demuxlength)['pb_500']` \text{ bp}$ after trimming) and moderate error rate (mean $\epsilon= `r format(readd(demuxqual)['pb_500'], digits = 4)`$ based on ccs quality scores) for the long amplicon in this study, the expected number of reads required to achieve two error-free reads is `r format(as.integer(round(2/(1 - readd(demuxqual)['pb_500'])^readd(demuxlength)['pb_500'])), big.mark = ",")`.
Given the high diversity relative to sequencing depth in this study (`r readd(region_table)$ASVs[readd(region_table)$seq_run == "pb_483" & readd(region_table)$region == "short"]` ASVs based on PacBio short amplicons, `r format(readd(demuxcounts)["pb_500"], format = "d", big.mark = ",")` trimmed long amplicon reads), this requirement could not have been met for the long amplicons except by the most abundant sequences.
In comparison, the equivalent requirement for the short amplicon ($L=`r readd(demuxlength)['pb_483']` \text{ bp}$, $\epsilon=`r readd(demuxqual)['pb_483']`$) is only `r round(2/(1 - readd(demuxqual)['pb_483'])^readd(demuxlength)['pb_483'], 1)` reads.

We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into shorter homologous domains, independently denoising each segment, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
We used this method, detailed below, for all of the PacBio and Ion Torrent datasets.
The method is not applicable to Illumina paired-end reads.

Raw reads were divided into shorter regions by matching to covariance models (CM), which are similar to stochastic hidden markov models (HMM), but account for both nucleotide sequence and RNA secondary structure [@eddy1994].
First, the 5.8S rDNA was located in each read by searching for Rfam model RF0002 [@kalvari2018] using `cmsearch` from Infernal 1.1.2 [@nawrocki2013], and all bases before the 5.8S were assigned to ITS1.
No attempt was made to remove the approximately 12\ bp fragment of the SSU from the 5' end of ITS1 in the long amplicons; it was too short to be reliably detected by a CM or the HMMs employed by ITSx [@bengtsson-palme2013].
<!--This approach was able to identify both the 5' and 3' ends of the 5.8S rDNA more consistently and quickly than ITSx [@bengtsson-palme2013, data not shown].-->
A reference alignment including conserved RNA base pairing between and within the 5.8S and relevant portions of LSU was generated from the fungal 28S RNA seed alignment from RDP release 11.5 [@glockner2017; @cole2014] by truncating after the LR5 primer site and using the reference line to annotate the variable regions *sensu* @michot1984 and @raue1988.
A CM was generated from the alignment using `cmbuild` from Infernal.
The fragment of each read beginning with the 5.8S rDNA was then aligned to the CM using `cmalign` from Infernal.
The annotation line in the CM alignment for each read was then used to split the reads into alternating more-conserved and less-conserved domains as shown in Figure\ \@ref(fig:regions),
where LSU1-4 represent the conserved domains of LSU flanking the variable D1-3 domains [@michot1984].
For short amplicons, only (partial) 5.8S, ITS2, and (partial) LSU1 were extracted.
Code to extract the segments, including annotated seed alignments and CMs, is available in the new R package `LSUx`, available on Github at https://github.com/brendanf/LSUx.

Each of the extracted domains was independently filtered for length (Supplementary Table\ \@ref(tab:region-limits)) and a maximum of three expected errors, and denoised using DADA2.
To test the performance of DADA2 on intermediate-length amplicons, adjacent domains from each read were concatenated to form full ITS (ITS1 + 5.8S + ITS2), partial LSU (LSU1 + D1 + LSU2 + D2 + LSU3 + D3 + LSU4), and partial 32S (5.8S + ITS2 + LSU) datasets, and these were also denoised. 
The DADA2 error model was fit using the 5.8S region for long amplicons, and using the entire read for short amplicons.
Independent error models were fit for each sequencing strategy (i.e., long *vs.* short amplicons, different sequencing technologies).
DADA2 was run with complete pooling for PacBio libraries, and pseudo-pooling
for Ion Torrent libraries.
Chimeras within each region were removed using `removeBimeraDenovoTable` from DADA2.

For each raw read, the denoised sequences of the different domains were concatenated to form a set of full-length sequences.
For reads which were not assigned a denoised sequence for each domain, the raw read for the domain was used instead.
Because ITS2 is the most variable of the amplified domains as indicated by the greatest number of unique ASVs (Figure\ \@ref(fig:regions)b), reads with identical ITS2 regions are expected to have highly similar sequences in the other regions, unless the amplicon was chimeric.
Therefore, concatenated sequences with identical ITS2 were clustered, and each cluster was aligned in R using the `DECIPHER` package [@wright2015].
Outlier sequences, as determined by mean pairwise distance from the rest of the alignment, were removed from each alignment using the `odseq` package [@jehl2015], using the default threshold of 0.025.
The consensus of the remaining aligned sequences was assigned as the full-length ASV sequence for the entire cluster.
Full-length ASV sequences with more than three ambiguous bases (i.e., no nucleotide >50% at a given position) were removed.
The count and sample distribution of reads assigned to each full-length ASV were calculated in order to form a sample × ASV community matrix.
A similar process was used to generate a consensus ITS (ITS1--5.8S--ITS2) and LSU (LSU1--D1--LSU2--D2--LSU3--D3--LSU4) sequence for each ASV.
The process of assigning consensus full-length ASVs was carried out using the new `tzara` package for R, available at https://github.com/brendanf/tzara.

@@

Because the Illumina dataset consisted of paired-end reads, domains could not be extracted prior to denoising, and the ASVs generated according to the standard DADA2 workflow were used.
The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the results from the other technologies.

@@

To assess the completeness of sequencing using each of the sequencing strategies, ASV and OTU accumulation curves were generated using the iNEXT package [citation]. Additionally, average estimated ASV and OTU richness was calculated for each sample after rarefying to 100 sequences using 100 repetitions. ASV and OTU richness were then compared for each biological sample as sequenced by the different strategies. For each pair of strategies, the best fit slope for the ASV richness recovered by each strategy was calculated by total least squares using the "deming" package [citation]. For this analysis, samples in the PacBio and Illumina datasets were combined bioinformatically to match the corresponding samples from the incompletely demultiplexed Ion Torrent dataset.

The effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, was determined by PERMANOVA [@anderson2003] using the `adonis2` function in the R package `vegan` [@R-vegan].
Comparisons were made between the fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina short, PacBio short, PacBio long).

In order to detect bias at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(readd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(readd(tech_class_table))` classes), were included.
The community matrix was normalized to proportional abundance by dividing the counts for each class in each sample by the total number of reads for that sample, but were not rarefied to even sampling depth [@mcmurdie2014].
PERMANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs. short); and sequencing technology (Illumina MiSeq vs. PacBio RS II).
Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the `capscale` function in `vegan` [@R-vegan].
Spatiotemporal effects were partialled out in order to visualize effects due to sequencing technology and amplicon length.

@@

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline differed between sequencing strategies (Table\ \@ref(tab:bioinfo)).
Sequencing with PacBio RS II yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads (<= 1 expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \@ref(fig:qual-check)b).
Although the per-base read quality of the long amplicon PacBio sequences was similar to that of Illumina (Figure \@ref(fig:qual-check)a), this tranlated to a greater number of expected errors per read due to the amplicon length (Figure \@ref(fig:qual-check)b).
Demultiplexing, primer trimming, and quality filtering reduced the read totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long amplicons, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
In contrast, extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads, `r percent(1-readd(filtercounts_ITS2)['pb_483']/readd(rawcounts)['pb_483'], format = "d")` of trimmed short amplicon PacBio reads, and `r percent(1-readd(filtercounts_ITS2)['is_057']/readd(rawcounts)['is_057'], format = "d")` of trimmed Ion Torrent reads.
This represented greater loss of PacBio short amplicon reads, but less loss of PacBio long amplicon reads and Ion Torrent reads.

Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Figure\ \@ref(fig:full-length)a).
Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Figure\ \@ref(fig:full-length)b).
The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r readd(length_kw)$statistic %>% formatC(format = "e", digits = 2)`, $p < 2.2\times10^{-16}$),
but the difference between means was fairly small, with mean amplicon lengths of `r filter(readd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
In contrast, the length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(readd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.

The length distribution of the different regions extracted from the long amplicon are shown in Figure\ \@ref(fig:region-lengths).
ITS1 showed the greatest length variability (mean $\pm$ standard deviation: `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp), followed by ITS2 (`r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp) and the variable regions in LSU (D2: `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D3: `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D1: `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp).
Approximately `r (readd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500", length > 200) %$% sum(reads)/readd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500") %$% sum(reads)) %>% signif(1) %>% percent()` of reads included an intron of 40--60 bp in the LSU4 region, not visible in Figure\ \@ref(fig:region-lengths) due to rarity.
Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations < 2 bp.

@@

<!-- ADD HERE results about the alpha diversity comparison -->

@@

Taxonomic composition of the sequenced soil fungal community at the class level is is summarized in Figure\ \@ref(fig:fungi-comm) and as a heat tree [@foster2017] in Figure\ \@ref(fig:taxa).
The ML tree for fungal ASVs, along with taxonomic assignments, is shown in Supplementary File 3.
According to the PHYLOTAX assignments, Fungi represented `r percent(readd(fungi_asv)["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(readd(fungi_read)["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(readd(fungi_asv)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(fungi_read)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm) and \@ref(fig:ppcoa)).

Fungi categorized as ECM made up `r percent(readd(ecm_asv)["pb_500"], format = "f", digits = 1)` of ASVs and `r percent(readd(ecm_read)["pb_500"], format = "f", digits = 1)` of reads in the long amplicon library, and `r paste(percent(range(readd(ecm_asv)[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(ecm_read)[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library (Figure\ \@ref(fig:ecm)).
Although amplicon length had a significant effect on ECM community composition at the family level, the explained variation was very low (PERMANOVA with 9999 permutations, $p=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), and the majority of variation was again spatiotemporal ($p<`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$).
Variation between sequencing technologies was not significant ($p=`r formatC(readd(tech_ecm_fam_adonis)["tech", "p.value"], format = "f", digits = 2)`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["tech", "R2"], format = "f", digits = 4)`$).

@@

@@

Both datasets from the short amplicon library showed a puzzling pattern with no autocorrelation at 0\ m and 1\ year, but a weak negative correlation at 1\ m and 1\ year.
The general shape of the correlograms were similar for long and short amplicon datasets.
We hypothesize that this pattern may result from a superposition of positively and negatively correlated temporal processes operating at different spatial scales.
As an example, the positive year-to-year correlation at 0\ m distance could be due to one species or set of species which strongly persist between years, but are primarily confined to patch sizes $<1$\ m,
while the negative year-to-year correlation could be due to a different species or set of species which deplete nutrients within a $>1$\ m patch over the course of a year, and are likely to be absent from the same patch in the following year.

>>>>

base.dir <- regmatches(getwd(), regexpr(".*oueme-fungi-transect", getwd()))
write_bib(c("base", "phyloseq", "vegan", "metacoder", "deming"), file = "R.bib", tweak = TRUE)

@@

Because of the variety of sequencing platforms and analytical pipelines which have been used in metabarcoding studies, comparisons between studies may be difficult.
Here we investigated the effects of different sequencing strategies and post-analysis on biological conclusions using measurement of the spatiotemporal turnover rate of the fungal community in an ECM-dominated woodland in Benin by metabarcoding of bulk soil, sampled at narrow intervals, over two years.
Turnover scale is the distance at which two communities can be considered to be independent samples of the local species pool.
Knowledge of turnover scale is important when planning studies of local diversity and its environmental correlates.
Turnover scale varies between different ecosystems and taxonomic groups.
Turnover scale is often measured by the range at which a Mantel correlogram indicates significant autocorrelation, or by fitting a function to an empirical distance-decay curve of community dissimilarity vs. distance [@legendre2012].

@@

We conclude that this poor performance was due to a combination of long amplicon length and low sequencing depth relative to community diversity (see Supplementary Methods).
We therefore developed a new workflow to assemble ASVs from the long amplicons by splitting the reads into homologous domains (Figure\ \@ref(fig:regions)a), independently denoising each segment, concatenating the denoised domains, and finally forming consensus sequences based on ITS2 identity.
We used this method, implemented in the new R packages `LSUx` (https://github.com/brendanf/LSUx) and `tzara` (https://github.com/brendanf/tzara) and detailed in the Supplementary Methods, for all of the PacBio and Ion Torrent datasets.

@@

Because the `LSUx` plus `tzara` method as currently implemented is not applicable to Illumina paired-end reads, the ASVs generated from the Illumina dataset according to the standard DADA2 workflow were used.
The ITS2 region was extracted from the ASVs using `LSUx` for comparison to the results from the other technologies.

@@

To assess the completeness of sequencing using each of the sequencing strategies, ASV and OTU accumulation curves were generated using the `iNEXT` package [@hsieh2016]. Additionally, average estimated ASV and OTU richness was calculated for each sample after rarefying to 100 sequences using 100 repetitions. ASV and OTU richness were then compared for each biological sample as sequenced by the different strategies. For each pair of strategies, the best fit slope for the ASV richness recovered by each strategy was calculated by total least squares using the `deming` package [@R-deming]. For this analysis, samples in the PacBio and Illumina datasets were combined bioinformatically to match the corresponding samples from the incompletely demultiplexed Ion Torrent dataset.

The effect of sequencing technology and amplicon length on the recovered fungal community composition, as assessed by the Bray-Curtis dissimilarity, was determined by PERMANOVA [@anderson2003] using the `adonis2` function in the `vegan` package [@R-vegan].
Comparisons were made between the fungal communities recovered by the three sequencing strategies that were successfully demultiplexed (Illumina short, PacBio short, PacBio long).

In order to detect potential biases at larger taxonomic scales, ASVs were clustered according to the assigned taxonomic class.
Only samples where all three strategies yielded at least 100\ fungal reads (`r phyloseq::nsamples(readd(tech_class_table))/3` samples), and classes which represented at least 1% of reads in at least one sample (`r phyloseq::ntaxa(readd(tech_class_table))` classes), were included.
The community matrix was normalized to proportional abundance by dividing the counts for each class in each sample by the total number of reads for that sample, but were not rarefied to even sampling depth [@mcmurdie2014].
PERMANOVA included three terms: an indicator for soil sample, comprising all spatiotemporal effects; amplicon length (long vs. short); and sequencing technology (Illumina MiSeq vs. PacBio RS II).
Partial Principal Coordinates Analysis (PPCoA) was applied to the same dissimilarity matrix using the `capscale` function in `vegan` [@R-vegan].
Spatiotemporal effects, i.e. differences between different samples, were partialed out using a `Condition()` term in the `capscale` model in order to visualize effects due to sequencing technology and amplicon length.

@@

The number of sequencing reads and ASVs at each stage in the bioinformatics pipeline differed between sequencing strategies (Table\ \@ref(tab:bioinfo)).
Sequencing with PacBio RS II yielded more than twice as many raw reads for long amplicons as for short amplicons, with approximately 125 thousand and 50 thousand reads, respectively.
Ion Torrent Ion S5 and Illumina MiSeq yielded substantially more reads, with 20.7 million and 10.8 million, respectively.
PacBio sequencing of the short amplicon library yielded the highest fraction of high-quality reads ($\le 1$ expected error), followed by Illumina, with Ion Torrent yielding the lowest quality (Figure \@ref(fig:qual-check)b).
Although the per-base read quality of the long amplicon PacBio sequences was similar to that of Illumina (Figure \@ref(fig:qual-check)a), this tranlated to a greater number of expected errors per read due to the amplicon length (Figure \@ref(fig:qual-check)b).
Demultiplexing, primer trimming, and quality filtering reduced the read totals by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_500"] %>% percent(format = "d")` for PacBio long amplicons, but only by `r (1-readd(filtercounts_full)/readd(rawcounts))["pb_483"] %>% percent(format = "d")` for PacBio short amplicons, resulting in a similar number of filtered reads for the two strategies.
Losses in demultiplexing, trimming, and quality filtering were intermediate for Ion Torrent and Illumina, with `r (1-readd(filtercounts_full)/readd(rawcounts))[c("is_057", "SH-2257")] %>% percent(format = "d") %>% text_list()` loss, respectively.
In contrast, extraction of only the ITS2 region before quality filtering (Figure \@ref(fig:qual-check)c) resulted in the loss of `r percent(1-readd(filtercounts_ITS2)['pb_500']/readd(rawcounts)['pb_500'], format = "d")` of trimmed long amplicon PacBio reads.

Almost all of the short amplicons from all three technologies were between 240 and 375\ bp long (Figure\ \@ref(fig:full-length)a).
Although the length profile of the three sequencing runs were similar, Illumina MiSeq had the largest fraction of reads near the top of the range, followed by Ion Torrent Ion S5 and PacBio RS II (Figure\ \@ref(fig:full-length)b).
The difference in length distributions was statistically significant due to the large sample size (Kruskal-Wallis statistic = `r readd(length_kw)$statistic %>% formatC(format = "e", digits = 2)`, $p < 2.2\times10^{-16}$),
but the difference between means was fairly small, with mean amplicon lengths of `r filter(readd(reads_table), region == "short") %>% group_by(seq_run) %>% summarize(length = weighted.mean(length, reads)) %>% deframe() %>% magrittr::extract(c("pb_483", "is_057", "SH-2257")) %>% text_list(digits = 0)` bp for PacBio, Ion Torrent, and Illumina, respectively.
The length of the long amplicon reads varied widely, from `r filter(readd(reads_table), region == "long") %$% range(length) %>% paste(collapse = " to ")` bp, with a mean of `r filter(readd(reads_table), region == "long") %$% weighted.mean(length, reads) %>% round()` bp.

Among the different regions extracted from the long amplicon (Figure\ \@ref(fig:region-lengths),
ITS1 showed the greatest length variability (mean $\pm$ standard deviation: `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp), followed by ITS2 (`r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "ITS2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp) and the variable regions in LSU (D2: `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D2", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D3: `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D3", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp; D1: `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.mean(length, weight = reads) %>% round()` $\pm$ `r readd(reads_table) %>% filter(region == "D1", seq_run == "pb_500") %$% reldist::wtd.var(length, weight = reads) %>% sqrt() %>% round()`\ bp).
Approximately `r (readd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500", length > 200) %$% sum(reads)/readd(reads_table) %>% filter(region == "LSU4", seq_run == "pb_500") %$% sum(reads)) %>% signif(1) %>% percent()` of reads included an intron of 40--60 bp in the LSU4 region, not visible in Figure\ \@ref(fig:region-lengths) due to rarity.
Except for these sequences, all conserved regions of LSU, as well as 5.8S, displayed very little size variation, as expected, with standard deviations < 2 bp.
A larger fraction of ITS2 sequences extracted from the long amplicon dataset were shorter than 140 bp than from the short amplicon datasets (Figure \@ref(fig:ITS2-length)).
The taxonomic identity of these sequences will be discussed below.

@@

ASV and OTU richness estimates after rarefaction were strongly correlated between the three sequencing technologies applied to the short amplicon library ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Short"), endsWith(as.character(strategy_y), "Short"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
The slope of the relationship between PacBio and Illumina richness estimates was not significantly different from 1, indicating that these two technologies give essentially identical rarefied richness estimates, dispite the approximately 200× difference in original sequencing depth.
Ion Torrent resulted in rarefied richness estimates which were `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), startsWith(as.character(strategy_x), "Ion"), endsWith(as.character(strategy_y), "Short"))$value_x %>% divide_by(1, .) %>% subtract(1) %>% range() %>% percent(format = "f", digits = 0) %>% paste(collapse = " to ")` greater than the other technologies, an effect which is also visible in ASV (Figure\ \@ref(fig:accum-ASV)) and OTU (Figure\ \@ref(fig:accum-OTU)) accumulation curves.
Richness estimates were somewhat less strongly correlated between the PacBio long amplicon dataset and the three short amplicon datasets ($R^2 =$ `r dplyr::filter(dplyr::bind_rows(readd(demingfits_ASV), readd(demingfits_OTU)), endsWith(as.character(strategy_x), "Long") | endsWith(as.character(strategy_y), "Long"))$r_squared %>% range() %>% format(format = "f", digits = 2) %>% paste(collapse = " to ")`; Figure\ \@ref(fig:alpha-compare)).
Although the total least squares regression lines indicate that the long amplicon dataset resulted in richness estimates which were more similar to the short amplicon results from Ion Torrent than the other two technologies, the confidence intervals for all slope estimates overlap with each other and 1.

@@

Taxonomic composition of the sequenced soil fungal community at the class level is is summarized in Figure\ \@ref(fig:fungi-comm) and as a heat tree [@foster2017] in Figure\ \@ref(fig:taxa).
The ML tree for fungal ASVs, along with taxonomic assignments, is shown in Supplementary File 3.
According to the PHYLOTAX assignments, Fungi represented `r percent(readd(fungi_asv)["PacBio_Long"], format = "f", digits = 0)` of the ASVs and `r percent(readd(fungi_read)["PacBio_Long"], format = "f", digits = 0)` of the reads in the long amplicon library, compared to `r paste(percent(range(readd(fungi_asv)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(fungi_read)[c("PacBio_Short", "Illumina_Short", "Ion.Torrent_Short")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library.
Many of the ASVs which were unique to the long-amplicon library thus fall outside kingdom Fungi (Figure\ \@ref(fig:heattree-length-compare)).
In particular a large fraction of ITS2 sequences with length less than 140 (Figure\ \@ref(fig:ITS2-length)) were identified as Alveolates (Figure\ \@ref(fig:short-its)).

Measured fungal community composition at the class level varied significantly between long and short amplicons (PERMANOVA with 9999 permutations, $p<`r formatC(readd(tech_class_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), but only marginally between sequencing technologies ($p=`r formatC(readd(tech_class_adonis)["tech", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["tech", "R2"], format = "f", digits = 3)`$).
The majority of variation was spatiotemporal (i.e., between samples; $p<`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_class_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$), but once this variation was removed, the remaining effect consisted of a clear bias against Sordariomycetes in the long amplicon dataset (Figures\ \@ref(fig:fungi-comm) and \@ref(fig:ppcoa)).

Fungi categorized as ECM made up `r percent(readd(ecm_asv)["pb_500"], format = "f", digits = 1)` of ASVs and `r percent(readd(ecm_read)["pb_500"], format = "f", digits = 1)` of reads in the long amplicon library, and `r paste(percent(range(readd(ecm_asv)[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the ASVs and `r paste(percent(range(readd(ecm_read)[c("pb_483", "SH-2257", "is_057")]), format = "f", digits = 1), collapse = "–")` of the reads in the short amplicon library (Figure\ \@ref(fig:ecm)).
Although amplicon length had a significant effect on ECM community composition at the family level (Figure\ \@ref(fig:ecm-heattree)), the explained variation was very low (PERMANOVA with 9999 permutations, $p=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["amplicon", "R2"], format = "f", digits = 3)`$), and the majority of variation was again spatiotemporal ($p<`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "p.value"], format = "f")`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["paste(site, x, year)", "R2"], format = "f", digits = 2)`$).
Variation between sequencing technologies was not significant ($p=`r formatC(readd(tech_ecm_fam_adonis)["tech", "p.value"], format = "f", digits = 2)`$, $R^2=`r formatC(readd(tech_ecm_fam_adonis)["tech", "R2"], format = "f", digits = 4)`$).

@@

<!-- 592 I’d start the discussion with a couple of introductory sentences pointing out the main finding and guiding the reader to the next sections.-->

@@

Both datasets from the short amplicon library showed no temporal autocorrelation, but a weak negative spatiotemporal correlation at 1\ m and 1\ year, which was probably a statistical artifact.
The correlograms for the long amplicon datasets showed similar trends.

Weighted UniFrac did not reliably detect spatial structure within this relatively ecologically homogeneous community.
Although the Mantel test did show a small but significant positive autocorrelation in the fungal community at the smallest size category (1\ m; Figure\ \@ref(fig:correlog)), the distance-decay plot in Figure\ \@ref(fig:spatio) does not show any clear relationship.
The functional fit showed poor convergence, with a 95% confidence interval for spatial range of 0--5700\ m, indicating little evidence of spatial structure.
This is probably because the majority of community turnover in this system, especially among ECM fungi, is between closely related species or individuals of the same species, while the presence of major clades (e.g., ECM lineages *sensu* @Tedersoo2010) are more spatially constant.
This is also reflected in the generally smaller sample-to-sample dissimilarities measured by UniFrac as compared to Bray-Curtis in Figure \@ref(fig:spatio).
UniFrac would be more suited at larger spatial scales and/or larger ecological gradients.
